# HW2

## Question
3.4, 3.11, and 3.20


## Answer

### 3.4 
The Rayleigh density [156, Ch. 18] is
$$
f(x)=\frac{x}{\sigma^{2}} e^{-x^{2} /\left(2 \sigma^{2}\right)}, \quad x \geq 0, \sigma>0 .
$$
Develop an algorithm to generate random samples from a Rayleigh $(\sigma)$ distribution. Generate Rayleigh $(\sigma)$ samples for several choices of $\sigma>0$ and check that the mode of the generated samples is close to the theoretical mode $\sigma$ (check the histogram).

*解:* 

Rayleigh随机变量$X$的分布函数:
$$
F(x)=1-\exp \left(-\frac{x^{2}}{2 \sigma^{2}}\right), x \geq 0
$$
所以$F^{-1}(y)=\sigma \sqrt{-2 \ln (1-y)}$

```{r}
Rayleigh =function(sigma, n){
  for(i in 1:n) {
    U=runif(n)
    V=1-U
    X = sigma * sqrt(-2 * log(V))
  }
  return(X)
}
sigma = 2
n = 1000

hist(Rayleigh(sigma, n),main = "Rayleigh",xlab="")
```

#### 多试几组sigma
```{r}
sigma=c(1:9)
n=1000
par(mfrow=c(3,3))
for(i in 1:9){
  title=paste0("sigma=",sigma[i])
  hist(Rayleigh(sigma[i], n),main = title,xlab="")
}
```





### 3.11
Generate a random sample of size 1000 from a normal location mixture. The components of the mixture have $N(0,1)$ and $N(3,1)$ distributions with mixing probabilities $p_{1}$ and $p_{2}=1-p_{1}$. Graph the histogram of the sample with density superimposed, for $p_{1}=0.75 .$ Repeat with different values for $p_{1}$ and observe whether the empirical distribution of the mixture appears to be bimodal. Make a conjecture about the values of $p_{1}$ that produce bimodal mixtures.

 
#### p1取0.75
```{r}
n=1000
X1=rnorm(n,0,1)
X2=rnorm(n,3,1)
u <- runif(n)
p1=as.integer(u < 0.75) 
p2=1-p1
Z=p1*X1+p2*X2
hist(Z,main = "p1=0.75")
```


#### p1从0变化到1
```{r}
p=seq(0,1,0.1)
par(mfrow=c(3,4))
for(i in 1:11){
  p1=as.integer(u < p[i]) 
  p2=1-p1
  Z=p1*X1+p2*X2
  title=paste0("p1=",p[i])
  hist(Z,main=title)
}
```


把$p1$的值从0变化到1，发现: 当$p1$在$0.5$附近的时候容易出现双峰.


### 3.20
A compound Poisson process is a stochastic process $\{X(t), t \geq 0\}$ that can be represented as the random sum $X(t)=\sum_{i=1}^{N(t)} Y_{i}, t \geq 0$, where $\{N(t), t \geq 0\}$ is a Poisson process and $Y_{1}, Y_{2}, \ldots$ are id and independent of $\{N(t), t \geq 0\}$. Write a program to simulate a compound Poisson $(\lambda)$-Gamma process $(Y$ has a Gamma distribution). Estimate the mean and the variance of $X(10)$ for several choices of the parameters and compare with the theoretical values. Hint: Show that $E[X(t)]=\lambda t E\left[Y_{1}\right]$ and $\operatorname{Var}(X(t))=\lambda t E\left[Y_{1}^{2}\right]$.



```{r}
Poisson_Gamma=function(n, t, lambda, r, beta) {
  N =rpois(n, lambda * t)
  X=sapply(N, function(N, r, beta) sum(rgamma(N, r,beta)), r,beta)
  return(X)
}

test=function(n, t, lambda, r, beta) {
  x=Poisson_Gamma(n, t, lambda, r,beta)
  ## 样本均值
  sm=mean(x) 
  ## 理论均值
  vm=var(x)
  ## 样本方差
  tm=lambda * t * r/beta
  ## 理论方差
  tv=lambda * t * (1 + r) * r/beta^2
  
  ## 输出结果样式
  cat("r=",r,"beta=",beta,"\n")
  cat("样本均值:", sm, "  ")
  cat("理论均值:", vm, "\n")
  cat("样本方差:", tm, "  ")
  cat("理论方差:", tv, "\n\n")
}

## 参数值
n=1000
lambda_seq=c(1:3)
r_seq=c(1:3)
beta_seq=c(1:3)
t=10

for (lambda in lambda_seq) {
  for (r in r_seq) {
    for (beta in beta_seq) {
      test(n, t, lambda,r,beta)
    }
  }
}

```






